Near and Mid-infrared properties of known $z\geq5$ Quasars

### 2019-06-12 tex/VHzQ.tex
VHzQ update

\begin{figure}
  \centering
  \includegraphics[width=8.5cm]
  {../light_curves/MIR_LCs/NEOWISER_LC_histogramlog_20180827.png}
  \vspace{-16pt}
  \caption[]
  {Histogram showing the total number of exposures for each detected VH$z$Q
    from the full combination of ALLWISE, NEOWISE and NEOWISE-R.} 
  \label{fig:MIR_LC_epochs}
\end{figure}
\subsection{Variability}
VH$z$Qs, if accreting at, or above the Eddington limit, might well have changing mass accretion rates, $\dot{m}_{\rm accr}$. A consequence of this would be that these quasars exhibit signs of variability, most likely showing up in their UV/optical rest-frame emission. We look for evidence of this variability signature in the NIR and MIR light-curves of the VH$z$Qs.

Quasars are known to have dramatically changing Balmer lines, especially H$\beta$ \citep[e.g.,][]{LaMassa2015, Ruan2016, Runnoe2016, Macleod2016, Gezari2017, Runco2016, YangQ2018, Assef2018, Stern2018, Ross2018, MacLeod2019, Graham2019}. As a guide, we note that H$\alpha$ is redshifted to 3.94$\mu$m (i.e. W1) at $z=5.00$ and 5.57$\mu$m, which is between the W2 and W3-bands at $z=7.50$. 
H$\beta$ is redshifted to 2.92$\mu$m, which is the blue edge of W1, at $z=5.00$ and 
4.13$\mu$m at $z=7.50$. 
%%
Less well understood is the temporal behaviour of the metal lines, in particular \civ and \mgii. 
\civ enters the
$Y$-band at redshift $z$=5.32 and exits at $z$=5.99, and enters the
$J$-band at redshift $z=6.55$ and exits at $z$=7.57. \mgii enters the
$H$-band at redshift $z=4.33$ and exits at $z=5.37$ and enters the
$K$-band at redshift $z=6.25$ and exits at $7.50$.

\begin{figure}
  \includegraphics[width=8.5cm]
  {/cos_pc19a_npr/programs/quasars/highest_z/light_curves/MIR_LCs/three_MIR_LC_egs_20180827.pdf}
  \centering
  \caption[]
  {Here we show the MIR NEOWISE-R for J0100+2802 \citep{Wu2015}, J0224-4711 and  J1626+2751. 
    Red points are the W1 band; cyan points the W2 band.} 
  \label{fig:MIR_LC_3egs}
\end{figure}

Figure~\ref{fig:MIR_LC_epochs} gives the number of NEOWISE-R epochs
and detections there are for each VH$z$Q, while
Figure~\ref{fig:MIR_LC_3egs} presents three examples of the MIR
lightcurves and associated colour changes. Here we show J0100+2802
\citep{Wu2015}, J0224-4711 and J1626+2751.

Using the extended datasets described in Section~\ref{sec:NIR_data}, 
we select objects with at least 8 measurements observed with intervals of 
at least 30 days, in at least one filter. 
The signal-to-noise on the average calibrated flux is required to be $\geq8$. 
%(i.e. {\tt aperJky3}$>0$. 
%%(i.e. the average calibrated flux over all epochs must be >0) 
%and 
%{\tt aperJky3/aperJky3Err}$>=$8)
%%
A requirement of SNR$>3$ per `average' epoch, where `average epoch' is
dependent on the averaging timescales is also required.
The clipped median and standard deviation is then calculated 
\begin{equation}
      var = 1.48 \times  {\rm m.a.d.} / \bar{\epsilon}
\end{equation}
where $var$ is the index of variation, 
m.a.d. is the median absolute deivation, 
and $\bar{\epsilon}$ is the mean of the error in each point in the
light curve, divided by the total number of points.
%%
With these criteria, there are 21 WFCAM and 12 VISTA objects 
with 1 object (VIK J1148+0056 ) in common to both, and this sample of 32 is our starting point for variability investigations.

We present four objects that are particularly well-sampled in the NIR; these are
MMTJ0215-0529 (Figure~\ref{fig:MMTJ0215-0529}), 
CFHQSJ0216-0455 (Figure~\ref{fig:CFHQSJ0216-0455}), 
SHELLQsJ0220-0432 (Figure~\ref{fig:SHELLQsJ0220-0432}) and 
SDSS J0959+0227 (Figure~\ref{fig:SDSS J0959+0227}). 
We also incorporate the MIR NEOWISE-R light data where available.

\begin{figure}
  \begin{subfigure}{}
    \centering
    \includegraphics[width=8.5cm]{../light_curves/MMTJ0215-0529LC_20190214.png}
    \caption{MMT J0215-0529}
    \label{fig:MMTJ0215-0529}
  \end{subfigure} 
  \begin{subfigure}{}
    \centering
    \includegraphics[width=8.5cm]{../light_curves/CFHQSJ0216-0455LC_20190214.png}
    \caption{CFHQS J0216-0455}
    \label{fig:CFHQSJ0216-0455}
  \end{subfigure}
\end{figure}

@@

Longer time-series, better SNR and moreover, $(1+z)$. 
Doing this independently in each band...

>>>>

\begin{figure}
  \centering
  \includegraphics[width=8.5cm]
  {../light_curves/MIR_LCs/NEOWISER_LC_histogramlog_20180827.png}
  \vspace{-16pt}
  \caption[]
  {Histogram showing the total number of exposures for each detected VH$z$Q
    from the full combination of ALLWISE, NEOWISE and NEOWISE-R.} 
  \label{fig:MIR_LC_epochs}
\end{figure}
\subsection{Variability}
VH$z$Qs, if accreting at, or above the Eddington limit, might well
have changing mass accretion rates, i.e., $\ddot{m}_{\rm accr}$. A
consequence of this would be that these quasars exhibit signs of
variability, most likely showing up in their UV/optical rest-frame
emission. We look for evidence of this variability signature in the
NIR and MIR light-curves of the VH$z$Qs.

Quasars are known to have dramatically changing Balmer lines,
especially H$\beta$ \citep[e.g.,][]{LaMassa2015, Ruan2016, Runnoe2016,
Macleod2016, Gezari2017, Runco2016, YangQ2018, Assef2018, Stern2018,
Ross2018, MacLeod2019, Graham2019}. As a guide, we note that H$\alpha$
is redshifted to 3.94$\mu$m (i.e. W1) at $z=5.00$ and 5.57$\mu$m,
which is between the W2 and W3-bands at $z=7.50$.  H$\beta$ is
redshifted to 2.92$\mu$m, which is the blue edge of W1, at $z=5.00$
and 4.13$\mu$m at $z=7.50$. Less well understood is the temporal
behaviour of the metal lines, in particular \civ and \mgii.  \civ
enters the $Y$-band at redshift $z$=5.32 and exits at $z$=5.99, and
enters the $J$-band at redshift $z=6.55$ and exits at $z$=7.57. \mgii
enters the $H$-band at redshift $z=4.33$ and exits at $z=5.37$ and
enters the $K$-band at redshift $z=6.25$ and exits at $7.50$.

\begin{figure}
  \includegraphics[width=8.5cm]
  {/cos_pc19a_npr/programs/quasars/highest_z/light_curves/MIR_LCs/three_MIR_LC_egs_20180827.pdf}
  \centering
  \caption[]
  {Here we show the MIR NEOWISE-R for J0100+2802 \citep{Wu2015}, J0224-4711 and  J1626+2751. 
    Red points are the W1 band; cyan points the W2 band.} 
  \label{fig:MIR_LC_3egs}
\end{figure}
Figure~\ref{fig:MIR_LC_epochs} gives the number of NEOWISE-R epochs
and detections there are for each VH$z$Q, while
Figure~\ref{fig:MIR_LC_3egs} presents three examples of the MIR
lightcurves and associated colour changes. Here we show J0100+2802
\citep{Wu2015}, J0224-4711 and J1626+2751.

Using the extended datasets described in Section~\ref{sec:NIR_data}, 
we select objects with at least 8 measurements observed with intervals of 
at least 30 days, in at least one filter. 
The the average calibrated flux over all epochs must be $>0$ 
(i.e. {\tt aperJky3}$>0$) 
with the signal-to-noise is required to be $\geq8$. 
(i.e. {\tt aperJky3/aperJky3Err}$>=$8).
The clipped median and standard deviation is then calculated 
\begin{equation}
      var = 1.48 \times  {\rm m.a.d.} / \bar{\epsilon}
\end{equation}
where $var$ is the index of variation, m.a.d. is the median absolute
deivation, and $\bar{\epsilon}$ is the mean of the error in each point
in the light curve, divided by the total number of points.  With these
criteria, there are 21 WFCAM and 12 VISTA objects with 1 object (VIK
J1148+0056) in common to both, and this sample of 32 is our starting
point for variability investigations.

We present four objects that are particularly well-sampled in the NIR. 
These are: 
MMTJ0215-0529 (Figure~\ref{fig:MMTJ0215-0529}), 
CFHQSJ0216-0455 (Figure~\ref{fig:CFHQSJ0216-0455}), 
SHELLQsJ0220-0432 (Figure~\ref{fig:SHELLQsJ0220-0432}) and 
SDSS J0959+0227 (Figure~\ref{fig:SDSSJ0959+0227}). 
We incorporate the MIR NEOWISE-R light curve data where available.

\begin{figure}
  \begin{subfigure}{}
    \centering
    \includegraphics[width=8.5cm]{../light_curves/MMTJ0215-0529LC_20190214.png}
    \caption{The infrared light curve for MMT J0215-0529 (}
    \label{fig:MMTJ0215-0529}
  \end{subfigure} 
  \begin{subfigure}{}
    \centering
    \includegraphics[width=8.5cm]{../light_curves/CFHQSJ0216-0455LC_20190214.png}
    \caption{CFHQS J0216-0455}
    \label{fig:CFHQSJ0216-0455}
  \end{subfigure}
\end{figure}

@@

Even with well sampled data across the 3000-4000 day observed time-scales that the WFCAM and VISTA surveys span, the $(1+z)$ time-dilation dramatically affects the sampled rest-frame timescales sampled.

SDSS J0959+0227 is in the VH$z$Q sample from \citet{Wang2016} 
%% Wang et al. 2016, ApJ, 819, 24, and their Table1. 
but is first reported in \citet{Civano2011} as CID-2220 (CXOCJ095908.1+022707), 
%2011, ApJ, 741, 91 
a high-redshift, $z>3$ AGN in the Chandra-COSMOS field. 
%%
The spectrum from this object is presented in \citet{Ikeda2012} 
% et al., 2012, ApJ, 756, 160, 
and is called object ``No. 11'' in that papers' Table 1. 
and it is a beauty!! Figure 2 with the spectrum is attached. 
\citet{Ikeda2012} show SDSS J0959+0227 having a narrow Ly$\alpha$ line, 
and would likely be a Lyman-$\alpha$ emitting galaxy had it not been an X-ray source, 
The X-ray luminosity is $\approx3x10^44$ erg s$-1$ in the 2–10 keV rest-frame, 
so it is an AGN. This object is clearly not a regular broadline ``Type 1'' AGN, but  
if there is even the slightest of variability, this could potentially be a high-$z$ 
AGN transitioning from a narrow-line ``Type 2'' object to a broadline Type 1 quasar.

Doing this independently in each band...
